\documentclass[defaultstyle,11pt]{thesis}

\usepackage{amssymb}		% to get all AMS symbols
\usepackage{amsmath}
\usepackage{multirow}
\usepackage[table,xcdraw]{xcolor}
%\usepackage{xcolor}
%\usepackage{subfloat}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}		% to insert figures
\usepackage{hyperref}		% PDF hyperreferences??

%https://tex.stackexchange.com/questions/3450/make-pdf-open-in-fill-window-mode-with-hyperref
%XYZ, Fit, FitH, FitV, FitR, FitB, FitBH, FitBV
\hypersetup{pdfstartview={FitH null null 1.00}}

%%%%%%%%%%%%   All the preamble material:   %%%%%%%%%%%%

\title{Sum Over Histories Representation of Chemical Kinetics: An Interpretive and Predictive Method for Modeling Chemical Kinetics Using Time-dependent Pathways}

\author{Shirong}{Bai}

\otherdegrees{B.S., University of Science and Technology of China, 2012 \\
	      M.S., University of Colorado Boulder, 2016}

\degree{Doctor of Philosophy, Chemical Physics}		%  #1 {long descr.}
	{Ph.D., Chemical Physics}		%  #2 {short descr.}

\dept{Department of}			%  #1 {designation}
	{Chemistry and Biochemistry}		%  #2 {name}

\advisor{Prof.}				%  #1 {title}
	{Rex T. Skodje}			%  #2 {name}

\reader{Prof.~Robert Parson}		%  2nd person to sign thesis
%\readerThree{Ms.~Thora Nea}		%  3rd person to sign thesis

\abstract{  %\OnePageChapter	% because it is very short
%Chemical kinetics can be viewed as the cooperative propagation of chemical %substances. This idea leads us to use a chemical network to study the cooperative behavior of chemical species. What's more, because of the changing nature of chemical systems, a time-dependent chemical network was adopted in the work of this thesis.
Chemical kinetics can be viewed as an intricate network of inter-related chemical reactions that work cooperatively to convert reagent species into product species. The network is in general time dependent reflecting the non-steady state nature of the chemistry. When it comes to interpreting and predicting chemical kinetics, the history of chemical moieties can play vital roles. In order to study the histories of chemical substances using time-dependent chemical network, this thesis focuses on developing a Sum Over Histories Representation (SOHR for short) of chemical kinetics. The description of time-dependent chemistry of a reaction network is provided by chemical pathways defined at a molecular level. Using this methodology, the quantitative time evolution of the kinetics is described by enumerating the most important pathways followed by a chemical moiety such as a tagged atom. An explicit formula for the pathway probabilities is derived which takes the form of an integral over a time-ordered product. This expression has a simple and computationally efficient Monte Carlo representation which permits the method to be applied to a wide range of problems.
\newline
\paragraph{}
In SOHR, the history of the chemical moiety can be described by time-dependent pathways. Unlike the static flux methods for path analysis, SOHR includes the explicit time-dependence of pathway probabilities. Using SOHR, the sensitivity of an observable with respect to a kinetic parameter such as a rate coefficient is then analyzed in terms of how that parameter affects the chemical pathway probabilities. This thesis demonstrates that large sensitives are often associated with rate limiting steps along important chemical pathways or by reactions that control the branching of reactive flux, though they vary with time.
\newline
\paragraph{}
In addition to interpreting chemical kinetics, this thesis studies the practical approach to modeling chemical kinetics without solving conventional mass-action ODEs. An iterative framework was introduced that allows the time-dependent pathway probabilities to be generated from a knowledge of elementary rate coefficients. To avoid the pitfall of integrating over the histories of long paths, we proposed a sector-by-sector strategy that shortens the candidate path without losing numerical accuracy. This method was successfully applied to the model Lotka-Volterra system and to a realistic H$_2$ combustion system.
\newline
\paragraph{}    
This thesis culminates with a discussion of the interpretative and predictive applicability of SOHR.




	}

\dedication[Dedication]{	% NEVER use \OnePageChapter here.
	\begin{center}
	To Zhangying Bai
	\end{center}
	}

\acknowledgements{	% \OnePageChapter	% *MUST* BE ONLY ONE PAGE!
To my dismay, I discovered that it is impossible at the end of my PhD career to recall all of the occasions when family, mentors, friends, acquaintances and strangers gave me invaluable advice, criticism and instructions. The following is in no way intended to be an exclusive list.
\newline
\paragraph{} 
First and foremost, I would like to thank Rex T. Skodje for being knowledgeable, courageous and creative as a mentor and being endlessly supportive as a friend. Along the path you supervised me, it took me years to appreciate your wisdom and great depth of thought, to which I can only hope to emulate. I am so grateful for the freedom you gave me regarding work schedule and research topics selection. I will fondly remember the good weeks we spent conversing and developing unique research ideas every day, though I was constantly reminded of “nothing is new under the sun”. As a friend, your encouragement and indulgence lit my path as well. I have enjoyed the sunny and windless summers we worked at Argonne National laboratory, Dalian Institute of Chemical Physics, and CU Boulder. Most importantly, I will forever bear in heart your good jokes. I have also been so lucky to collaborate with Dr. Michael Davis and Dr. Raghu Sivaramakrishnan, I tremendously benefited from your expertise. I am also grateful to my undergraduate advisor Prof. Zhonghuai Hou who taught me the power of hard work. I am also grateful to Prof. Xueming Yang who demonstrated that a great scientist can be a warm person as well. Finally, thank you to Prof. Veronica Vaida and Prof. Robert Parson for always saying yes to my requests.
\newline
\paragraph{} 
Closer to home, I would like to thank Jonathan Church and Robert Wells for helping me improve my oral English and being my editor-in-chief. I am grateful to Dr. Premashis Manna; your patience, elegance, and permanent smile have swept away the darkness of graduate school for an international student. I am also indebted to Ye Yuan, Xinshuo Yang, Xiaohang Zhang, Yuqing Xia and Huaxing Sun. The wisdom you shared when we were having lunch, playing video games, playing basketball, playing soccer, or playing tennis encouraged me and inspired me.
\newline
\paragraph{} 
I am also appreciative of my many other mentors and teachers who have shaped me. I am deeply grateful to my middle school teacher Yaqin Zhu who believed I had a talent in math, my middle school teacher Chaoliang Yang who first whet my appetite of poetry, my high school teacher Youqing Ding who showed me the beauty of Chemistry, and my high school teacher Shaopei Wei who ignited my interest in Physics. The people without whom this thesis would certainly not have been possible are my grandmother, my grandfather, my little brother and my parents: five of the kindest, smartest, most honest and most diligent people I know.

	}

% \IRBprotocol{E927F29.001X}	% optional!

\ToCisShort	% use this only for 1-page Table of Contents

\LoFisShort	% use this only for 1-page Table of Figures
% \emptyLoF	% use this if there is no List of Figures

\LoTisShort	% use this only for 1-page Table of Tables
% \emptyLoT	% use this if there is no List of Tables

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%       BEGIN DOCUMENT...         %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\input macros.tex
\input chapter1.tex
\input chapter2.tex
\input chapter2dot5.tex
\input chapter3.tex
\input chapter4.tex
\input chapter5.tex

%%%%%%%%%   then the Bibliography, if any   %%%%%%%%%
\bibliographystyle{plain}	% or "plain", "siam", or "alpha", or, "unsrt", etc.
\nocite{*}		% list all refs in database, cited or not
\bibliography{refs}		% Bib database in "refs.bib"

%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
\appendix
\input appendixA.tex


\end{document}

